### 2014-02-12 paper/manuscript.tex
Made all figures PDFs, and formatted them for publishing.
Some final changes to the manuscript.
- also prepping the repo for publication, with changes to the
  readme, dodo file, etc.

\graphicspath{{../figures/bw/}}
%% \newcommand{\hidefigures}{\relax}

@@

Noise and heterogeneity are both known to benefit neural coding.
Stochastic resonance describes how noise,
in the form of random fluctuations in a neuron's membrane voltage,
can improve neural representations of an input signal.
Neuronal heterogeneity refers to variation in any one of a number of neuron parameters,
and is also known to increase the information content of a population.
We explore the interaction between noise and heterogeneity
and find that their benefits to neural coding are not independent.
Specifically, a neuronal population better represents an input signal
when either noise or heterogeneity is added,
but adding both does not always improve representation further.
To explain this phenomenon, we propose
that noise and heterogeneity both operate using two shared mechanisms:
1) temporally desynchronizing the firing of neurons in the population,
and 2) linearizing the response of a population to a stimulus.
We first characterize the effects of noise and heterogeneity
on the information content of populations of
either leaky integrate-and-fire or FitzHugh-Nagumo neurons.
We then examine how the mechanisms
of desynchronization and linearization produce these effects,
and find that they work to distribute information equally across all neurons in the population,
both in terms of signal timing (desynchronization) and in terms of signal amplitude (linearization).
Without noise or heterogeneity,
all neurons in the population encode the same aspects of the input signal;
adding noise or heterogeneity allows neurons
to encode \emph{complementary} aspects of the input signal,
thereby increasing information content.
The simulations detailed in this paper
highlight the importance of heterogeneity and noise in population coding,
demonstrate their complex interactions in terms of the information content of neurons,
and explain these effects in terms of underlying mechanisms.

@@

A new form of stochastic resonance---called \emph{suprathreshold} stochastic resonance (SSR)---has emerged in the past decade \citep{McDonnell2009,McDonnell2011}. In neuroscience, suprathreshold SR looks at the benefits of noise to populations of nonlinear neurons receiving a common signal of which a significant portion is above the systems' firing threshold(s). This concept was first introduced by \cite{Stocks2000}, who looked at groups of binary threshold units and found that noise increases information transmission in such systems. Subsequent research has examined the SSR effect in populations of more sophisticated simulated neurons \citep{Stocks2001}.

@@

We performed numerical experiments analyzing how noise and heterogeneity affect the information capacity of populations of both FitzHugh-Nagumo (FHN) neurons and leaky integrate-and-fire (LIF) neurons. Heterogeneity was added to a population by selecting the firing thresholds ${b_i}$ from a uniform distribution $\mathcal{U}(-b_r, b_r)$, where $b_r$ controls the degree of heterogeneity, with larger $b_r$ corresponding to a more heterogeneous population (on average). Noise was added to the voltage parameter of the neuron model via a stochastic offset $\eta(t)$, Gaussian distributed with zero-mean and standard deviation $\sigma_\eta$, and independent and identically distributed for each neuron. For the remainder of this paper, we will use the term ``noise'' to refer specifically to this stochastic offset. At each level of heterogeneity and noise, 100 trials were performed to achieve an accurate estimate of the mean information content, with each trial using a unique set of random biases and a unique set of stochastic offsets added to the membrane voltage.

@@

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics{experiment.eps}
  \fi
  \caption{
    \textbf{The experimental setup.} In our experiment, neurons are fed a common aperiodic input. Each neuron also has an independent white noise process and a randomly chosen bias that are added to the input. Varying the amplitude of the noise and the width of the interval from which the biases are randomly drawn allows us to control the amount of noise and heterogeneity in the experiment, respectively. Additionally, some ``off'' neurons invert their input (not shown). The output of each FHN or LIF neuron model is summed and filtered to produce the population output; this is compared with the input signal using mutual information as a metric.
  }
  \label{fig:exp}
\end{figure}

@@

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/infohetero.eps}
  \fi
  \caption{
    \textbf{Heterogeneity increases information content of a population of neurons, and exhibits a resonance effect.}
    Adding some heterogeneity (x-axis) to a population of LIF or FHN neurons increases their ability to represent an input signal (y-axis), while too much heterogeneity decreases representational ability. This effect is only observed at lower noise levels; at higher levels of noise, the noise alone is able to fully desynchronize neuronal firing and linearize neuron responses, so adding heterogeneity offers no additional benefit.
  }
  \label{fig:infohetero}
\end{figure}

@@

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/infocontour.eps}
  \fi
  \caption{
    \textbf{The benefits of noise and heterogeneity to information representation are not independent.}
We measured the information content (top, in bits) and information rate (bottom, in bits per spike),
across many levels of noise and heterogeneity.
In LIF neurons, both noise and heterogeneity show independent resonance effects:
given that only one of noise or heterogeneity is present,
there is a non-zero and finite level of noise or heterogeneity that optimizes information content.
Setting both noise and heterogeneity to their (independent) optimal levels
fails to further improve the representational ability of the population,
showing that these effects are not independent.
We propose that both parameters use
the same underlying mechanisms of desynchronization and linearization,
so once the population has been desynchronized and linearized by one parameter,
adding the other cannot further improve information content.
Heterogeneity is unable to fully desynchronize FHN neurons (see \textsc{\nameref{scn:desync}}),
so information content in these neurons is maximized for a specific combination of heterogeneity and noise. However, this effect is not seen when examining the information rate,
indicating that noise benefits neural systems partly through increasing firing rates.
  }
  \label{fig:infocontour}
\end{figure}

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/infonoise.eps}
  \fi
  \caption{
    \textbf{Noise increases information content of homogeneous populations of FHN and LIF neurons.} A) In homogeneous populations, noise exhibits a resonance effect on information content. B) The resonance effect of noise is less pronounced when measuring the information rate, indicating that noise operates partially by increasing firing rates.
  }
  \label{fig:infonoise}
\end{figure}

@@

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/phase.eps}
  \fi
  \caption{
    \textbf{Noise and heterogeneity both desynchronize neurons.} To measure whether noise and heterogeneity have a desynchronizing effect on populations of neurons, we measured the standard deviation of neuron phases across the population (see \textsc{\nameref{scn:methods}}), where a smaller standard deviation indicates increased synchronization. Heterogeneity is seen to significantly desynchronize both FHN and LIF neurons, specifically across low to moderate levels of noise. At high levels of noise, the noise itself is sufficient to desynchronize neurons, and additional heterogeneity has no effect.
  }
  \label{fig:phase}
\end{figure}

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/syncraster.eps}
  \fi
  \caption{
    \textbf{The effects of noise and heterogeneity on LIF neuron firing.}
    Spike rasters illustrate the effects of noise and heterogeneity parameters on neuronal firing in a population of 32 LIF neurons. With low noise and heterogeneity, neuronal firing is highly synchronized. Adding a moderate amount of noise or heterogeneity desynchronizes this firing. Adding a moderate level of both parameters cannot provide further desynchronization, since firing is already quite desynchronized. High levels of noise cause rapid firing in almost all neurons, degrading the ability of the population to represent the input signal. High levels of heterogeneity cause the signal to be entirely above or below the firing thresholds of many neurons, again degrading representational ability. Noise: $\sigma_\eta = $$10^{-3}$ (low), $10^{-2}$ (moderate), $10^{-1}$ (high). Heterogeneity: $b_r = $$10^{-2}$ (low), $10^{-1}$ (moderate), $10^{0}$ (high).
  }
  \label{fig:syncraster}
\end{figure}

@@

Similar to noise, heterogeneity acts to desynchronize neurons \citep{Burton2012}. Adding heterogeneity to a population increases the range of distribution of neuron phases (Figure~\ref{fig:phase}), thereby desynchronizing populations of both FHN and LIF neurons (Figure~\ref{fig:syncraster}). In a homogeneous population, changes in the input signal cross the common firing threshold of all neurons at the same time, resulting in synchronous firing. Contrast this with a heterogeneous population, where changes in the input signal sequentially cross the thresholds of neurons, causing some to begin firing before others. For example, an increasing input signal will cause neurons with lower thresholds to begin firing before those with higher thresholds, resulting in desynchronized firing within the population of neurons.

High frequency input signals may cross the thresholds of many neurons in a heterogeneous population at the same time, causing a large group of neurons to begin firing simultaneously. LIF neurons are still able to desynchronize quickly after such a stimulus, since the LIF tuning curves used have different firing rates for different sizes of suprathreshold signal (Figure~\ref{fig:tuning}). Specifically, a high-frequency increase in signal amplitude starts neurons firing at approximately the same time, but starts them firing at different rates, allowing them to quickly desynchronize. FHN neurons, however, all have around the same firing rate regardless of the strength of the suprathreshold signal (Figure~\ref{fig:tuning}), and heterogeneity with higher-frequency signals has a reduced effect, since these signals cause many neurons to begin firing at once, unable to desynchronize due to their common firing rate. This discrepancy between LIF and FHN neurons results in heterogeneity having a large effect on phase in LIF neurons (Figure~\ref{fig:phase}A), whereas the desynchronizing effect of heterogeneity in FHN neurons is more modest (Figure~\ref{fig:phase}B).

@@

Intrinsic noise makes neuron tuning curves more linear;
this phenomenon underlies many results in the stochastic resonance literature
\citep{Chialvo1997}.
Noise allows a neuron to respond to input signals below its firing threshold,
by occasionally raising the neuron's membrane voltage high enough
to cross the threshold and trigger an action potential.
This results in a non-zero firing rate for subthreshold signals,
with the firing rate proportional to the strength of the signal.
In type~II neuron models (\emph{e.g.} the FHN model),
noise can also slow firing for signals slightly above the firing threshold,
by momentarily pushing the neuron away from its limit cycle and back to near its unstable equilibrium.
Overall, the addition of noise has the effect
of linearizing the response of the neuron near its firing threshold,
resulting in a smoother transition from the silent to tonic firing regimes (Figure~\ref{fig:tuning}).
This result has been previously observed and extensively studied \citep{Stocks1996,Chialvo1997,Chance2002,Prescott2002,Shu2003}.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/tuningnoisy.eps}
  \fi
  \caption{
    \textbf{Noise linearizes neuron tuning curves.}
    We measured the tuning (stimulus-response) curves of LIF and FHN neurons under various levels of noise.
    We observe that noise linearizes the neuron tuning curves,
    Specifically, noise eliminates the sharp threshold in the tuning curve by
    occasionally pushing the neuron above its firing threshold
    for stimuli that are below the threshold.
    Noise also increases neuron firing rates across some or all stimuli.
  }
  \label{fig:tuning}
\end{figure}

Neuron heterogeneity has a linearizing effect similar to adding noise to a population, but unlike the noisy case, where it made sense to talk about noise linearizing an individual neuron's tuning curve, here we will discuss how heterogeneity linearizes the \emph{population's} tuning curve. The population's tuning curve is the set of points connecting the input signal strength to the decoded output. Since our model uses a simple summing decoder, the population tuning curve is equal to the sum of the individual neuron tuning curves. For a population of neurons with uniformly varied random firing thresholds, the population tuning curve is much more linear than a single neuron's tuning curve, whereas in the homogeneous case the population's tuning curve is the same as the neurons' tuning curve (Figure~\ref{fig:tuninghetero}). The linearizing effect of heterogeneity thereby allows for better representation of an input signal. This result is related to the original idea from \cite{Hubel1962} that variation in neural receptive fields provides a better basis for representing an input signal, and reappears in the recent heterogeneity literature \citep{Angelo2012}.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/tuninghetero.eps}
  \fi
  \caption{
    \textbf{Heterogeneity linearizes the response of a neuronal population to a stimulus.} A) Tuning curves for a heterogeneous population of $N = 9$ neurons. B) The population response of the heterogeneous population (heavy solid line) is much closer to a linear response (heavy dot-dashed line) than a single LIF tuning curve (heavy dashed line), even though the heterogeneous thresholds are far from evenly spaced.
  }
  \label{fig:tuninghetero}
\end{figure}

@@

Desynchronization operates by affecting the temporal properties of neuronal firing, and is only observed in spiking neuron models, such as the FHN and LIF models examined in this paper. The output of these models depends not only on the the output firing rate for a given input signal, as determined by the neuron tuning curve, but also by the timing of individual spike events. When nonlinear dynamical models of neurons are examined in a stochastic resonance simulation, as in \cite{Stocks2001}, the results are a combination of the temporally-based desynchronizing and response-time effects of noise, as well as the rate-based linearizing effects of noise.  Desynchronized neurons provide a better temporal basis for representing a signal, since each neuron is able to fire independently and code for a different temporal component of the signal. Even mild correlations between neurons are known to impede a population's representational ability \citep{Zohary1994}.

Our experimental setup was motivated by stochastic resonance experiments such as \cite{Stocks2001}, as well as computational \citep{Brody2003} and experimental \citep{Padmanabhan2010} work looking at heterogeneity. All these experiments present a common input signal to a number of neurons, with added noise or heterogeneity. The common input signal may seem artificial, but it allows us to examine how the properties of individual neurons---including the sign of the input signal (``on'' and ``off'' neurons), the bias current, and the membrane voltage---affect the ability of the neuronal population to encode an input signal. These heterogeneities in the neurons can be viewed as modulating the input signal presented to each neuron, such that each neuron receives a unique input current. In the case of homogeneous populations, all neurons do receive the same input current. This is clearly artificial, and emphasizes why theoreticians should include heterogeneity when modeling neural behaviour.

@@

\textbf{FitzHugh-Nagumo (FHN) model.} The FHN model is a simplification of the Hodgkin-Huxley model \citep{FitzHugh1961}. We chose to use the standard formulation presented by \cite{Izhikevich2006}, given for the $i$-th neuron by two coupled ordinary differential equations:
\begin{equation}
  \begin{array}{rl}
    \dot{v_i} & = v_i - \frac{1}{3}v_i^3 - w_i + \beta_i + e_i s(t) + \eta_i(t) \vspace{1em}\\
    \dot{w_i} & = 0.08(v_i - 0.8w_i + 0.7),
  \end{array}
  \label{eqn:fhn}
\end{equation}
where $v_i(t)$ is a fast (voltage) variable, $w_i(t)$ is a slow (recovery) variable, $\beta_i$ is a constant bias voltage, $e_i$ is a simple encoder equal to 1 for ``on'' neurons and $-1$ for ``off'' neurons, $s(t)$ is an input signal common to all neurons, and $\eta_i(t)$ is a Gaussian white noise process with autocorrelation $\left<\eta_i(t)\eta_j(t + \tau)\right> = \sigma^2 \delta_{ij}\delta(\tau)$. We chose $\beta_i = 0.3216 - b_i$, with the offset 0.3216 chosen such that the bias $b_i$ corresponds to the threshold at which the neuron begins firing.

@@

One shortfall of this comparison method is that it does not account for the delay between the input and output signals, due to the non-zero reaction time of the neurons. In neural systems, some delay is to be expected, and we should not penalize our simulated networks for not being able to encode and decode an input signal instantaneously. The difficulty in comparing a delayed version of the input signal with the output signal is that it raises the question of what length of delay should be used. Each pair of noise and heterogeneity levels may conceivably have a unique optimal delay, and in order to not favor either noise or heterogeneity, the optimal delay (\emph{i.e.,} the delay that produces the highest mutual information) should be calculated and used at each pair of noise and heterogeneity levels. In the interest of simplicity, we decided to instead compare the input and output signals with no delay, as is consistent with previous experiments \citep{Stocks2001}.

@@

\textbf{Heterogeneity.} To introduce heterogeneity into the population, we varied the bias voltages of the neurons. In homogeneous populations, all neurons had bias voltages equal to zero, the mean of the input signal. Heterogeneous populations had neurons with uniformly randomly distributed biases chosen from a range centered around the mean. In particular, three ranges were used: $[-0.01,0.01]$, $[-0.05,0.05]$, and $[-0.2,0.2]$, corresponding to small, medium and large levels of heterogeneity. The ranges were chosen to be roughly logarithmically distributed, with the largest range corresponding to biases that covered two standard deviations of the input signal.

\textbf{``On'' and ``off'' neurons.} To help minimize subthreshold stochastic resonance effects from significantly contributing to our results, we used populations that contained an equal number of ``on'' and ``off'' neurons, \emph{i.e.,} neurons that both increase firing rates as the input signal increases, and increase firing rates as the signal decreases. In terms of our model, this means that half the neurons of each population had $e_i = 1$, and the other half had $e_i = -1$. This means that half the population represents the positive and negative parts of the signal, respectively, and only a small amount of information is gained by the subthreshold SR effect, which allows each half of the population to represent both positive and negative signal components.

>>>>

\graphicspath{{../figures/}}
%% \newcommand{\hidefigures}{\relax}

@@

Noise and heterogeneity are both known to benefit neural coding.
Stochastic resonance describes how noise,
in the form of random fluctuations in a neuron's membrane voltage,
can improve neural representations of an input signal.
Neuronal heterogeneity refers to variation in any one of a number of neuron parameters,
and is also known to increase the information content of a population.
We explore the interaction between noise and heterogeneity
and find that their benefits to neural coding are not independent.
Specifically, a neuronal population better represents an input signal
when either noise or heterogeneity is added,
but adding both does not always improve representation further.
To explain this phenomenon, we propose
that noise and heterogeneity operate using two shared mechanisms:
1) temporally desynchronizing the firing of neurons in the population,
and 2) linearizing the response of a population to a stimulus.
We first characterize the effects of noise and heterogeneity
on the information content of populations of
either leaky integrate-and-fire or FitzHugh-Nagumo neurons.
We then examine how the mechanisms
of desynchronization and linearization produce these effects,
and find that they work to distribute information equally across all neurons in the population,
both in terms of signal timing (desynchronization) and in terms of signal amplitude (linearization).
Without noise or heterogeneity,
all neurons encode the same aspects of the input signal;
adding noise or heterogeneity allows neurons
to encode \emph{complementary} aspects of the input signal,
thereby increasing information content.
The simulations detailed in this paper
highlight the importance of heterogeneity and noise in population coding,
demonstrate their complex interactions in terms of the information content of neurons,
and explain these effects in terms of underlying mechanisms.

@@

A new form of stochastic resonance---called \emph{suprathreshold} stochastic resonance (SSR)---has emerged in the past decade \citep{McDonnell2009,McDonnell2011}. In neuroscience, SSR looks at the benefits of noise to populations of nonlinear neurons receiving a common signal of which a significant portion is above the systems' firing threshold(s). This concept was first introduced by \cite{Stocks2000}, who looked at groups of binary threshold units and found that noise increases information transmission in such systems. Subsequent research has examined the SSR effect in populations of more sophisticated simulated neurons \citep{Stocks2001}.

@@

We performed numerical experiments analyzing how noise and heterogeneity affect the information capacity of populations of both FitzHugh-Nagumo (FHN) neurons and leaky integrate-and-fire (LIF) neurons. Heterogeneity was added to a population by selecting the firing thresholds ${b_i}$ from a uniform distribution $\mathcal{U}(-b_r, b_r)$, where $b_r$ controls the degree of heterogeneity, with larger $b_r$ corresponding to a more heterogeneous population (on average). Noise was added to the voltage parameter of the neuron model via a stochastic offset $\eta(t)$, Gaussian distributed with zero-mean and standard deviation $\sigma_\eta$, and independent and identically distributed for each neuron. For the remainder of this paper, we will use the term ``noise'' to refer specifically to this stochastic offset. At each level of heterogeneity and noise, 100 trials were performed to achieve an accurate estimate of the mean information content, with each trial using a unique set of random biases and a unique set of stochastic offsets added to the membrane voltages.

@@

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics{figure1_experiment.pdf}
  \fi
  \caption{
    \textbf{The experimental setup.} In our experiment, neurons are fed a common aperiodic input. Each neuron also has an independent white noise process and a randomly chosen bias that are added to the input. Varying the amplitude of the noise and the width of the interval from which the biases are randomly drawn allows us to control the amount of noise and heterogeneity in the experiment, respectively. Additionally, some ``off'' neurons invert their input (not shown). The output of each FHN or LIF neuron model is summed and filtered to produce the population output; this is compared with the input signal using mutual information as a metric.
  }
  \label{fig:exp}
\end{figure}

@@

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{figure2_infohetero.pdf}
  \fi
  \caption{
    \textbf{Heterogeneity increases information content of a population of neurons, and exhibits a resonance effect.}
    Adding some heterogeneity (x-axis) to a population of LIF or FHN neurons increases their ability to represent an input signal (y-axis), while too much heterogeneity decreases representational ability. This effect is only observed at lower noise levels; at higher levels of noise, the noise alone is able to fully desynchronize neuronal firing and linearize neuron responses, so adding heterogeneity offers no additional benefit.
  }
  \label{fig:infohetero}
\end{figure}

@@

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{figure3_infocontour.pdf}
  \fi
  \caption{
    \textbf{The benefits of noise and heterogeneity to information representation are not independent.}
We measured the information content (top, in bits) and information rate (bottom, in bits per spike),
across many levels of noise and heterogeneity.
In LIF neurons, both noise and heterogeneity show independent resonance effects:
given that only one of noise or heterogeneity is present,
there is a non-zero and finite level of noise or heterogeneity that optimizes information content.
Setting both noise and heterogeneity to their (independent) optimal levels
fails to further improve the representational ability of the population,
showing that these effects are not independent.
We propose that both parameters use
the same underlying mechanisms of desynchronization and linearization,
so once the population has been desynchronized and linearized by one parameter,
adding the other cannot further improve information content.
Heterogeneity is unable to fully desynchronize FHN neurons (see \textsc{\nameref{scn:desync}}),
so information content in these neurons is maximized for a specific combination of heterogeneity and noise. However, this effect is not seen when examining the information rate,
indicating that noise benefits neural systems partly through increasing firing rates.
  }
  \label{fig:infocontour}
\end{figure}

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{figure4_infonoise.pdf}
  \fi
  \caption{
    \textbf{Noise increases information content of homogeneous populations of FHN and LIF neurons.} A) In homogeneous populations, noise exhibits a resonance effect on information content. B) The resonance effect of noise is less pronounced when measuring the information rate, indicating that noise operates partially by increasing firing rates.
  }
  \label{fig:infonoise}
\end{figure}

@@

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{figure5_phase.pdf}
  \fi
  \caption{
    \textbf{Noise and heterogeneity both desynchronize neurons.} To measure whether noise and heterogeneity have a desynchronizing effect on populations of neurons, we measured the standard deviation of neuron phases across the population (see \textsc{\nameref{scn:methods}}), where a smaller standard deviation indicates increased synchronization. Heterogeneity (individual traces) is seen to significantly desynchronize both FHN and LIF neurons, specifically across low to moderate levels of noise. At high levels of noise, the noise itself is sufficient to desynchronize neurons, and additional heterogeneity has no effect.
  }
  \label{fig:phase}
\end{figure}

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{figure6_syncraster.pdf}
  \fi
  \caption{
    \textbf{The effects of noise and heterogeneity on LIF neuron firing.}
    Spike rasters illustrate the effects of noise and heterogeneity parameters on neuronal firing in a population of 32 LIF neurons. With low noise and heterogeneity, neuronal firing is highly synchronized. Adding a moderate amount of noise or heterogeneity desynchronizes this firing. Adding a moderate level of both parameters cannot provide further desynchronization, since firing is already quite desynchronized. High levels of noise cause rapid firing in almost all neurons, degrading the ability of the population to represent the input signal. High levels of heterogeneity cause the signal to be entirely above or below the firing thresholds of many neurons, again degrading representational ability. Noise: $\sigma_\eta = $$10^{-3}$ (low), $10^{-2}$ (moderate), $10^{-1}$ (high). Heterogeneity: $b_r = $$10^{-2}$ (low), $10^{-1}$ (moderate), $10^{0}$ (high).
  }
  \label{fig:syncraster}
\end{figure}

@@

Similar to noise,
heterogeneity acts to desynchronize neurons \citep{Burton2012}.
Adding heterogeneity to a population
increases the range of distribution of neuron phases (Figure~\ref{fig:phase}),
indicating desynchronized firing
for both FHN and LIF neurons (Figure~\ref{fig:syncraster}).
Homogeneous populations exhibit synchronized firing
because changes in the input signal cross the common firing threshold
of all neurons at the same time.
Contrast this with a heterogeneous population,
where changes in the input signal
sequentially cross the thresholds of neurons,
causing some to begin firing before others.
For example, an increasing input signal
will cause neurons with lower thresholds
to begin firing before those with higher thresholds,
resulting in desynchronized firing within the population of neurons.

High frequency input signals may cross the thresholds of many neurons
in a heterogeneous population in a short period of time,
causing a large group of neurons to begin firing simultaneously.
LIF neurons are still able to desynchronize quickly after such a stimulus,
since the LIF tuning curves used have different firing rates
for different sizes of suprathreshold signal (Figure~\ref{fig:tuning}).
Specifically, a high-frequency increase in signal amplitude
starts neurons firing at approximately the same time,
but starts them firing at different rates,
allowing them to quickly desynchronize.
FHN neurons, however, all have around the same firing rate
regardless of the strength
of the suprathreshold signal (Figure~\ref{fig:tuning}).
In the presence of higher-frequency input signals
heterogeneity has a reduced effect in these neurons,
since such input signals cause many neurons to begin firing at once,
and they are unable to desynchronize due to their common firing rate.
This discrepancy between LIF and FHN neurons results in heterogeneity
having a large effect on phase in LIF neurons (Figure~\ref{fig:phase}A),
whereas the desynchronizing effect of heterogeneity
in FHN neurons is more modest (Figure~\ref{fig:phase}B).

@@

Intrinsic noise makes neuron tuning curves more linear
(Figure~\ref{fig:tuning});
this phenomenon underlies many results in the stochastic resonance literature
\citep{Chialvo1997}.
Noise allows a neuron to respond to input signals below its firing threshold,
by occasionally raising the neuron's membrane voltage high enough
to cross the threshold and trigger an action potential.
This results in a non-zero firing rate for subthreshold signals,
with the firing rate proportional to the strength of the signal.
In type~II neuron models (\emph{e.g.} the FHN model),
noise can also slow firing for signals slightly above the firing threshold
by momentarily pushing the neuron away from its limit cycle
and back toward its weakly unstable equilibrium.
Overall, the addition of noise has the effect
of linearizing the response of the neuron near its firing threshold,
resulting in a smoother transition from the silent to tonic firing regimes.
This result has been previously observed and extensively studied \citep{Stocks1996,Chialvo1997,Chance2002,Prescott2002,Shu2003}.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{figure7_tuningnoisy.pdf}
  \fi
  \caption{
    \textbf{Noise linearizes neuron tuning curves.}
    We measured the tuning (stimulus-response) curves of LIF and FHN neurons under various levels of noise.
    We observe that noise linearizes the neuron tuning curves,
    Specifically, noise eliminates the sharp threshold in the tuning curve by
    occasionally pushing the neuron above its firing threshold
    for stimuli that are below the threshold.
    Noise also increases neuron firing rates across some or all stimuli.
  }
  \label{fig:tuning}
\end{figure}

Neuron heterogeneity has a linearizing effect similar to noise,
but unlike the noisy case,
where it made sense to talk about noise
linearizing an individual neuron's tuning curve,
here we will discuss how heterogeneity
linearizes the \emph{population's} tuning curve.
The population's tuning curve is the set of points
connecting the input signal strength to the decoded output.
Since our model uses a simple summing decoder,
the population tuning curve is equal to
the sum of the individual neuron tuning curves.
For a population of neurons with uniformly varied random firing thresholds,
the population tuning curve is much more linear
than a single neuron's tuning curve,
whereas in the homogeneous case the population's tuning curve
is the same as the neurons' tuning curve (Figure~\ref{fig:tuninghetero}).
The linearizing effect of heterogeneity
thereby allows for better representation of an input signal.
This result is related to the original idea from \cite{Hubel1962}
that variation in neural receptive fields
provides a better basis for representing an input signal,
and reappears in the recent heterogeneity literature \citep{Angelo2012}.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{figure8_tuninghetero.pdf}
  \fi
  \caption{
    \textbf{Heterogeneity linearizes the response of a neuronal population to a stimulus.} A) Tuning curves for a heterogeneous population of $N = 9$ neurons. B) The population response of the heterogeneous population (heavy solid line) is much closer to a linear response (heavy dot-dashed line) than a single LIF tuning curve (heavy dashed line), even though the heterogeneous thresholds are far from evenly spaced.
  }
  \label{fig:tuninghetero}
\end{figure}

@@

Desynchronization operates by affecting the temporal properties of neuronal firing, and is only observed in spiking neuron models such as the FHN and LIF models examined in this paper. The output of these models depends not only on the the output firing rate for a given input signal, as determined by the neuron tuning curve, but also by the timing of individual spike events. When nonlinear dynamical models of neurons are examined in a stochastic resonance simulation, as in \cite{Stocks2001}, the results are a combination of the temporally-based desynchronizing and response-time effects of noise, as well as the rate-based linearizing effects of noise.  Desynchronized neurons provide a better temporal basis for representing a signal, since each neuron is able to fire independently and code for a different temporal component of the signal. Even mild correlations between neurons are known to impede a population's representational ability \citep{Zohary1994}.

Our experimental setup was motivated by stochastic resonance experiments such as \cite{Stocks2001}, as well as computational \citep{Brody2003} and experimental \citep{Padmanabhan2010} work looking at heterogeneity. All these experiments present a common input signal to a number of neurons, with added noise or heterogeneity. The common input signal may seem artificial, but it allows us to examine how the properties of individual neurons---including the sign of the input signal (``on'' versus ``off'' neurons) and the bias current---affect the ability of the neuronal population to encode an input signal. These heterogeneities in the neurons can be viewed as modulating the input signal presented to each neuron, such that each neuron receives a unique input current. In the case of homogeneous populations, all neurons do receive the same input current. This is clearly artificial, and emphasizes why theoreticians should include heterogeneity when modeling neural behaviour.

@@

\textbf{FitzHugh-Nagumo (FHN) model.} The FHN model is a simplification of the Hodgkin-Huxley model \citep{FitzHugh1961}. We chose to use the standard formulation presented by \cite{Izhikevich2006}, given for the $i$-th neuron by two coupled ordinary differential equations:
\begin{equation}
  \begin{array}{rl}
    \dot{v_i} & = v_i - \frac{1}{3}v_i^3 - w_i + \beta_i + e_i s(t) + \eta_i(t) \vspace{1em}\\
    \dot{w_i} & = 0.08(v_i - 0.8w_i + 0.7),
  \end{array}
  \label{eqn:fhn}
\end{equation}
where $v_i(t)$ is a fast (voltage) variable, $w_i(t)$ is a slow (recovery) variable, $\beta_i$ is a constant bias voltage, $e_i$ is a simple encoder equal to 1 for ``on'' neurons and $-1$ for ``off'' neurons, $s(t)$ is an input signal common to all neurons, and $\eta_i(t)$ is a Gaussian white noise process with autocorrelation $\left<\eta_i(t)\eta_j(t + \tau)\right> = \sigma_\eta^2 \delta_{ij}\delta(\tau)$. We chose $\beta_i = 0.3216 - b_i$, with the offset 0.3216 chosen such that the bias $b_i$ corresponds to the threshold at which the neuron begins firing.

@@

One shortfall of this comparison method is that it does not account for the delay between the input and output signals, due to the non-zero reaction time of the neurons. In neural systems, some delay is to be expected, and we should not penalize our simulated networks for not being able to encode and decode an input signal instantaneously. The difficulty in comparing a delayed version of the input signal with the output signal is that it raises the question of what length of delay should be used. Each pair of noise and heterogeneity levels may conceivably have a unique optimal delay, and in order to not favor either noise or heterogeneity, the optimal delay (\emph{i.e.,} the delay that produces the highest mutual information) should be calculated and used at each pair of noise and heterogeneity levels. In the interest of simplicity, we decided to instead compare the input and output signals with no delay, as is consistent with previous experiments \citep{Stocks2001}. In preliminary experiments, accounting for delay resulted in only a small increase in information.

@@

\textbf{Heterogeneity.} To introduce heterogeneity into the population, we varied the bias voltages $b_i$ of the neurons. In homogeneous populations, all neurons had bias voltages equal to zero, the mean of the input signal. Heterogeneous populations had neurons with uniformly randomly distributed biases chosen from a range $\mathcal{U}(-b_r, b_r)$, where $b_r$ is the radius of heterogeneity. For the phase plot (Figure~\ref{fig:phase}), three ranges were used: $b_r = 0.01, 0.05, 0.2$, corresponding to small, medium and large levels of heterogeneity. The ranges were chosen to be roughly logarithmically distributed, with the largest range corresponding to biases that covered two standard deviations of the input signal.

\textbf{``On'' and ``off'' neurons.} To help minimize subthreshold stochastic resonance effects from significantly contributing to our results, we used populations that contained an equal number of ``on'' and ``off'' neurons, \emph{i.e.,} neurons that both increase firing rates as the input signal increases, and increase firing rates as the signal decreases. In terms of our model, this means that half the neurons of each population had $e_i = 1$, and the other half had $e_i = -1$. This means that half the population represents the positive and negative parts of the signal, respectively, and only a small amount of information is gained by the subthreshold SR effect which allows each half of the population to represent both positive and negative signal components.
